---
title: "Survey Project"
output: html_document
author: Karthick Krishna Balaji
date: '2022-07-01'
---

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
library(tidyr)

survey <- read.csv("Get_to_know_survey_2022.csv")
```

## Question 1

Create a new column called Birth_Month which has only the month portion of Birth_Month_Year.

```{r message=FALSE, warning=FALSE}
survey <- survey %>% 
  rowwise() %>%
  mutate(Birth_Month = str_split(Birth_Month_Year,",")[[1]][1])
```

## Question 2

Create a new column called Birth_Year which has only the year portion of Birth_Month_Year

```{r message=FALSE, warning=FALSE}
survey <- survey %>% 
  rowwise() %>%
  mutate(Birth_Year = paste(str_split(Birth_Month_Year,", ")[[1]][-1],collapse=" "))
```

## Question 3

Since the birth month and birth year were open ended fields, there are many inconsistencies in the data. Standardize the values for Birth_Month and Birth_Year.

```{r message=FALSE, warning=FALSE}

survey$Birth_Month <- survey$Birth_Month %>% 
  str_replace_all(c("12"="December","11"="November","10"="October","9"="September","8"="August","7"="July","6"="June","5"="May","4"="April","3"="March","2"="February","1"="January","Feb"="February","Februaryruary"="February","Setpember" = "September","Septemeber" = "September"))

survey$Birth_Year<-as.numeric(survey$Birth_Year)

survey <- survey%>%
  mutate_if(is.character, str_trim)
```

## Question 4

Show the five most frequent birth months, and the five least frequent birth months.

```{r message=FALSE, warning=FALSE}

survey %>%
  group_by(Birth_Month) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  head(5)

survey %>%
  group_by(Birth_Month) %>%
  summarise(frequency = n()) %>%
  arrange(frequency) %>%
  head(5)
```

## Question 5

Show the five most frequent birth years, and the five least frequent birth year.

```{r message=FALSE, warning=FALSE}
survey%>%
  group_by(Birth_Year)%>%
  filter(!is.na(Birth_Year))%>%
  summarise(Frequency= n())%>%
  arrange(desc(Frequency))%>%
  head(5)
survey%>%
  group_by(Birth_Year)%>%
  filter(!is.na(Birth_Year))%>%
  summarise(Frequency= n())%>%
  arrange(Frequency)%>%
  head(5)
```

## Question 6

The State column was also open ended, so there are many inconsistencies. Clean this column to eliminate any non-US state, and to standardize the values for State.

```{r message=FALSE, warning=FALSE}

survey <- survey %>%
  transform(survey,State=ifelse((State %in% state.name | State %in% state.abb), State, NA)) %>%
  transform(survey,State=ifelse((State %in% state.abb), state.name[match(State, state.abb)] ,State))

survey<-survey[,c(1:9)]
```

## Question 7

Show the five most frequent states, and the five least frequent state

```{r message=FALSE, warning=FALSE}
survey%>%
  group_by(State)%>%
  filter(!is.na(State))%>%
  summarise(Frequency= n())%>%
  arrange(desc(Frequency))%>%
  head(5)
survey%>%
  group_by(State)%>%
  filter(!is.na(State))%>%
  summarise(Frequency= n())%>%
  arrange(Frequency)%>%
  head(5)
```

## Question 8

Similar to State, the column Country needs to be standardized. Once you standardize it, show the five most frequent countries, and the five least frequent countries.

```{r message=FALSE, warning=FALSE}

survey <- survey %>%
  separate_rows(Country,sep = "/" ,convert = TRUE) 

survey <- survey %>%
  transform(survey,Country=ifelse(Country %in% c("USA","United States","United State","USA","US of A","US"), "United States of America",Country))

survey<-survey[,c(1:9)]

survey <- survey%>%
  mutate_if(is.character, str_trim)

survey%>%
  group_by(Country)%>%
  summarise(Frequency= n())%>%
  arrange(desc(Frequency))%>%
  head(5)
survey%>%
  group_by(Country)%>%
  summarise(Frequency= n())%>%
  arrange(Frequency)%>%
  head(5)
```

## Question 9

We want to clean the How_Hear_About_Program column by eliminating the number and any extra character portion in those responses (so we keep only the text portion). Show the five most frequent responses to this column.

```{r message=FALSE, warning=FALSE}
survey$How_Hear_About_Program <- str_replace_all(survey$How_Hear_About_Program, ("^[[:digit:]]. "), "")

survey <- survey%>%
  mutate_if(is.character, str_trim)

survey%>%
  group_by(How_Hear_About_Program)%>%
  summarise(Frequency=n())%>%
  arrange(desc(Frequency))%>%
  head(5)
```

## Question 10

We want to transform the data from wide to long on column Languages by duplicating the values of the other columns. You can use the comma as a separator for Languages. Show the five most frequent languages.

```{r message=FALSE, warning=FALSE}
survey <- survey %>%
  separate_rows(Languages,sep = ", ")

survey <- survey%>%
  mutate_if(is.character, str_trim)

survey%>%
  group_by(Languages)%>%
  summarise(Frequency=n())%>%
  arrange(desc(Frequency))%>%
  head(5)
```
